---
title: "Lab 1: Introduction to Quarto"
author: "Christopher Hawkins"
date: "09/26/24"
editor: source
format: 
  html:
    toc: true
    toc-depth: 2
    toc-expand: true
    number-sections: true
    number-depth: 2
execute:
    echo: true
    error: true
code-fold: true
theme: minty
warning: false
message: false
embed-resources: true
---

# Part Two: Exploring Quarto Documents

## Specifying Global Execute Options

Execution options specify how the R code in your Quarto document should be displayed. This [guide](https://quarto.org/docs/computations/execution-options.html) provides descriptions on the options you can specify in a document's execution.

**Question 1** -- Add your name and the date to the YAML.

**Question 2** -- Add an `execute` line to your YAML that includes the following options, (1) your source code is always output on the page, and (2) your document will render even if there are errors.

## Tooth Growth Dataset

```{r}
#| eval: false
#| echo: true
#| label: practicing-running-code

# Load a dataset into the R Environment.
data(ToothGrowth)

# Look at the summary of the variables in the dataset.
summary(ToothGrowth)
```

**Question 3** -- Let's spice this code up a bit. Delete `1 + 1` from the code chunk and paste in the code provided in the Lab 1 instructions.

### Check the Data Documentation

**Question 4** -- In your **console** (*not* in the Quarto document), type `?ToothGrowth` (or alternatively `help(ToothGrowth)`.Use the information that pops up in the *Help* pane in RStudio to fill in the blanks below.

*This dataset investigates the effect of Vitamin C on tooth growth in Guinea Pigs.*

*The two supplement delivery methods include OJ (orange juice) and VC (ascorbic acid).*

*`ToothGrowth` is a data frame with 60 observations and 3 variables.*

**Question 5** -- *Before* the code chunk, change the "Running the Provided Code" section header to a header that better describes the contents of this section (e.g., Tooth Growth Dataset).

## Creating a Plot

**Question 6** -- Replace the `plot(pressure)` code with the code provided in the Lab 1 instructions.

```{r}
#| echo: true
#| label: creating-a-plot
#| layout-ncol: 2
#| fig-align: "center"
#| fig-cap:
#| - "Tooth Length vs Delivery Method."
#| - "Tooth Length vs Dosage Amount."

library(tidyverse)

ggplot(data = ToothGrowth, 
       mapping = aes(x = supp, y = len)) +
  geom_boxplot() +
  labs(x = "Supplement", y = "Length of Teeth (mm)")

ToothGrowth %>% 
  mutate(dose = as.factor(dose)) %>% 
  ggplot(mapping = aes(x = dose, y = len)) +
  geom_boxplot() +
  labs(x = "Dose of Supplement (mg/day)", y = "Length of Teeth (mm)")
```

Now, run this code chunk! You should see two side-by-side boxplots, one plot comparing tooth length between the two supplement delivery methods and one plot comparing tooth length between the three different dosages uses.

**Question 7** -- Read over the [Quarto documentation on figure options](https://quarto.org/docs/authoring/figures.html#computations) and add an option (`#|`) to the above code chunk to make the figures print side-by-side (i.e., in one row) in the rendered HTML file.

**Question 8** -- Specify in the code chunk options that these boxplots should be **center aligned**.

**Question 9** -- Specify a **figure caption** for the boxplots (using a code-chunk option).

## Performing a t-test

Perform a two-sided independent t-test to compare the resultant tooth length between the two supplement methods.

**Question 10** -– Using the t.test() function, write code to carry out the analysis. You can assume unequal variances and a two-sided alternative.

**Question 11** -– Run your code chunk to obtain the output for your statistical test.

**Question 12** -– Create a numbered list containing:

Your conclusion (in the context of these data) based on the p-value.

An interpretation of the confidence interval (make sure to read what confidence level is used by default).

```{r}
#| eval: true
#| echo: true
#| label: Performing a t-test

# Create a variable to hold supplement 1 data
VC <- ToothGrowth[1:30,1]
# Create a variable to hold supplement 2 data
OJ <- ToothGrowth[31:60,1]

# Assume unequal variances
# Assume two-sided alternative
testResults <- t.test(x = VC, y = OJ, alternative = "two.sided", var.equal = FALSE)

# Display the test results
print(testResults)

# Create a list containing the conclusion based on the p-value and an interpretation of the confidence interval.

# conclusion based on p-value:

p_value <- sprintf("Fail to reject null hypothesis, since p-value = %0.2f ≥ \u03B1 = 0.05.", testResults$p.value)
# reference: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/sprintf
# reference: https://stackoverflow.com/questions/60219962/is-it-possible-to-use-greek-letters-in-r-in-strings

# conclusion based on confidence interval:

confidence <- sprintf("The difference in means of VC and OJ is between %0.2f and %0.2f with a confidence level of %0.2f.", testResults$conf.int[1], testResults$conf.int[2], attr(testResults$conf.int, "conf.level"))

result_list <- list(p_value, confidence)

# print the list
print(result_list)
```
